A Bayesian search to find \\high-mass black holes in LIGO data

### 2020-08-21 main.tex
adjust some missing citation keys

Since the 1970s, there has been an accumulation of evidence for stellar-mass and supermassive black holes. In 2019, the Event Horizon Telescope provided the first visual evidence of the supermassive black hole M87  \cite{m87photo}.  As of January 2020, LIGO, Virgo and IAS have confirmed more than a dozen binary black hole systems \cite{GWTC1, IAS0, IAS1, IAS2}. These various discoveries have firmly established the existence of stellar-mass black holes, supermassive black holes and binary black hole systems.  Interestingly, there is still no direct evidence for intermediate-mass black holes, the black holes that lie in between stellar-mass and supermassive black hole systems with masses between $10^2-10^6$ \msun. Thus, the existence of intermediate-mass black holes is still speculative. \\

If intermediate-mass black holes are present, the gravitational waves emitted from the merger of a binary black hole system with at least one intermediate-mass black hole (up to a mass of $400$ \msun) will be detectable by the ground-based gravitational-wave network. Gravitational waves from such systems should occur at a rate of $0-10\text{yr}^{-1}$ \cite{fregeau2006imbhbRatePrediction, mandel2008rates,rodriguez2015bbhRatePredictions}.  However, even after conducting a targeted matched-filter based search for gravitational waves from intermediate-mass black holes the largest total mass detected so far is approximately $80$ \msun \cite{salemi2019search, abbott2019gwtc}. A possible explanation for the absence of intermediate-mass events may be due to their misclassification as short-duration instrumental noise transients known as glitches \cite{glitch_in_fifth_ligo_run, bayeswave, improving_dq_in_early_runs, ligo_glitch_gw150914, pycbc_short_duration_transients, pe_with_glitch, blip_glitches}. These glitches can mimic astrophysical signals and hence decrease the significance of true gravitational wave events. \\

@@

Following \citet{bcr_paper}, the \bcr is an odds ratio of the coherent signal hypotheses $\mathcal{H_S}$ and the incoherent instrumental feature hypothesis $\mathcal{H_I}$ for a network of $D$ detectors. $\mathcal{H_I}$ states that each $i$th detector has either pure Gaussian noise $\mathcal{H_N}$ or a glitch $\mathcal{H_G}$. The \bcr is given by
\begin{equation}
BCR = \frac{\alpha Z^S}{\Pi^D_{i=1}[\beta Z^G_i + (1-\beta)Z^N_i]}\ ,
\end{equation}
where $Z^S$, $Z_{Gi}$ and $Z_{Ni}$ are the Bayesian evidences (marginalised likelihoods) for $\mathcal{H_S}$, $\mathcal{H_N}$, and $\mathcal{H_G}$. $\alpha$ and $\beta$, are the prior odds for obtaining a signal $\alpha=P(\mathcal{H_S})/P(\mathcal{H_I})$, the prior odds for obtaining a glitch $\beta=P(\mathcal{H_G})/P(\mathcal{H_I})$. As the rate of signal and glitches are unknown, these priors $\alpha$ and $\beta$ are tuned to maximize the \bcr distributions for background data (signal-free data) and simulated signals $\cite{BCR1}.

 provides a Bayesian odds ratio comparing the probability that the data contains an astrophysical signal versus the probability that the data contains an incoherent glitch or noise. Although an odds ratio, the \bcr can be used  as a frequentistic \textit{ranking statistic}, ranking the likelihood that a trigger is due to a gravitational-wave signal \cite{bcr_paper}. The \bcr as a ranking statistic is much more powerful than a matched-filter SNR as the \bcr incorporates numerous features that may distinguish gravitational waves from noise and demands that signals must be coherent while glitches must be incoherent amongst different detectors. \\

@@

The Bayesian coherence ratio (BCR), as seen in \citet{bcr_paper}, is the odds between the hypothesis that the data comprise a
coherent, compact binary coalescence signal in Gaussian noise (\(\mathcal{H}_S\)), and the hypothesis that they instead
comprise incoherent instrumental features (\(\mathcal{H}_I\))---meaning each detector has either a glitch in Gaussian noise
(\(\mathcal{H}_N\)) or pure Gaussian noise (\(\mathcal{H}_G\)). For a network of \(D\) detectors,

@@

\caption[BCR vs iFAR for O1]{BCR versus iFAR distributions from simulated signals (blue), hardware injections (green), GW events (stars) and background triggers (black), taken from \citet{bcr_paper}.}\label{fig:bcrIfar}
\end{figure}

>>>>

Since the 1970s, there has been an accumulation of evidence for stellar-mass and supermassive black holes. In 2019, the Event Horizon Telescope provided the first visual evidence of the supermassive black hole M87  \cite{m87photo}.  As of January 2020, LIGO, Virgo and IAS have confirmed more than a dozen binary black hole systems \cite{GWTC1, IAS0, IAS1, IAS2, pycbc_ogc_2}. These various discoveries have firmly established the existence of stellar-mass black holes, supermassive black holes and binary black hole systems.  Interestingly, there is still no direct evidence for intermediate-mass black holes, the black holes that lie in between stellar-mass and supermassive black hole systems with masses between $10^2-10^6$ \msun. Thus, the existence of intermediate-mass black holes is still speculative. \\

If intermediate-mass black holes are present, the gravitational waves emitted from the merger of a binary black hole system with at least one intermediate-mass black hole (up to a mass of $400$ \msun) will be detectable by the ground-based gravitational-wave network. Gravitational waves from such systems should occur at a rate of $0-10\text{yr}^{-1}$ \cite{fregeau2006imbhbRatePrediction, mandel2008rates,rodriguez2015bbhRatePredictions}.  However, even after conducting a targeted matched-filter based search for gravitational waves from intermediate-mass black holes the largest total mass detected so far is approximately $80$ \msun \cite{imbhbSearch2014, ligo_imbh_search, abbott2019gwtc}. A possible explanation for the absence of intermediate-mass events may be due to their misclassification as short-duration instrumental noise transients known as glitches \cite{glitch_in_fifth_ligo_run, bayeswave, improving_dq_in_early_runs, ligo_glitch_gw150914, pycbc_short_duration_transients, pe_with_glitch, blip_glitches}. These glitches can mimic astrophysical signals and hence decrease the significance of true gravitational wave events. \\

@@

Following \citet{BCR1}, the \bcr is an odds ratio of the coherent signal hypotheses $\mathcal{H_S}$ and the incoherent instrumental feature hypothesis $\mathcal{H_I}$ for a network of $D$ detectors. $\mathcal{H_I}$ states that each $i$th detector has either pure Gaussian noise $\mathcal{H_N}$ or a glitch $\mathcal{H_G}$. The \bcr is given by
\begin{equation}
BCR = \frac{\alpha Z^S}{\Pi^D_{i=1}[\beta Z^G_i + (1-\beta)Z^N_i]}\ ,
\end{equation}
where $Z^S$, $Z_{Gi}$ and $Z_{Ni}$ are the Bayesian evidences (marginalised likelihoods) for $\mathcal{H_S}$, $\mathcal{H_N}$, and $\mathcal{H_G}$. $\alpha$ and $\beta$, are the prior odds for obtaining a signal $\alpha=P(\mathcal{H_S})/P(\mathcal{H_I})$, the prior odds for obtaining a glitch $\beta=P(\mathcal{H_G})/P(\mathcal{H_I})$. As the rate of signal and glitches are unknown, these priors $\alpha$ and $\beta$ are tuned to maximise the \bcr distributions for background data (signal-free data) and simulated signals $\cite{BCR1}.

 provides a Bayesian odds ratio comparing the probability that the data contains an astrophysical signal versus the probability that the data contains an incoherent glitch or noise. Although an odds ratio, the \bcr can be used  as a frequentistic \textit{ranking statistic}, ranking the likelihood that a trigger is due to a gravitational-wave signal \cite{BCR}. The \bcr as a ranking statistic is much more powerful than a matched-filter SNR as the \bcr incorporates numerous features that may distinguish gravitational waves from noise and demands that signals must be coherent while glitches must be incoherent amongst different detectors. \\

@@

The Bayesian coherence ratio (BCR), as seen in \citet{BCR1}, is the odds between the hypothesis that the data comprise a
coherent, compact binary coalescence signal in Gaussian noise (\(\mathcal{H}_S\)), and the hypothesis that they instead
comprise incoherent instrumental features (\(\mathcal{H}_I\))---meaning each detector has either a glitch in Gaussian noise
(\(\mathcal{H}_N\)) or pure Gaussian noise (\(\mathcal{H}_G\)). For a network of \(D\) detectors,

@@

\caption[BCR vs iFAR for O1]{BCR versus iFAR distributions from simulated signals (blue), hardware injections (green), GW events (stars) and background triggers (black), taken from \citet{BCR1}.}\label{fig:bcrIfar}
\end{figure}
